---
title: "Random subset variable A"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r define_paths}
path_GLODAP_preprocessing <-
  paste(path_root, "/observations/preprocessing/", sep = "")
path_cmorized <-
  "/nfs/kryo/work/loher/CESM_output/RECCAP2/cmorized_Dec2020/split_monthly/"
path_preprocessing  <-
  paste(path_root, "/model/preprocessing/", sep = "")
```

```{r load_libraries_specific, include=FALSE}
library(stars)
library(lubridate)
library(rqdatatable)
library(gsw)
library(marelac)
```

# Distribution of GLODAP observations in the pre- and post-2010 era

```{r GLODAP_distribution_maps}

GLODAP <-
  read_csv(paste(path_GLODAP_preprocessing,
                 "GLODAPv2.2020_preprocessed.csv",
                 sep = ""))

section <- map + 
  scale_fill_viridis_c(direction = -1) +
      coord_quickmap(expand = 0) +
  guides()
  

a <- section +
      geom_bin2d(data = GLODAP %>% filter(year <= 2010),
                 aes(lon, lat),
                 binwidth = 1) +
      labs(
        subtitle = paste("Nr of observations:", nrow(GLODAP %>% filter(year <= 2010))),
        x = "Longitude",
        y = "Latitude"
      )

b <- section +
      geom_bin2d(data = GLODAP %>% filter(year > 2010),
                 aes(lon, lat),
                 binwidth = 1) +
      labs(
        subtitle = paste("Nr of observations:", nrow(GLODAP %>% filter(year > 2010))),
        x = "Longitude",
        y = "Latitude"
      )

a / b +
  plot_layout(guides = "collect")

```

# Randomly subset model data

Here we randomly subset cmorized (1x1) model with variable forcing, according to the total number of GLODAP observations for the whole period from a previously cleaned file. The number for the annual subset remains the same for each year, which could be expressed by the total number of observations divided by number of years.

Besides, Model results are given in [mol m^-3^], whereas GLODAP data are in [µmol kg^-1^]. This refers to the variables:

- DIC  
- ALK  
- O2  
- NO3  
- PO4  
- SiO3  
- AOU (calculated)

For comparison, model results were converted from [mol m^-3^] to [µmol kg^-1^]

# Spatial distribution

```{r random_resample_distribution_maps}

# read in random model sampling file
resample_lat <-
  read_csv(paste(path_preprocessing,
                 "GLODAPv2.2020_preprocessed_model_runA_random_subset_lat.csv",
                 sep = ""))

# plot random sampling cmorized grids in each year
#i_year <- "2000"
years <- c("1982", "1990", "2000", "2010", "2019")
for (i_year in years) {
  # i_year <- years[1]
  
  resample_lat_year <- resample_lat %>%
    filter(year == i_year)
  
  print(
    map +
      geom_bin2d(data = resample_lat_year,
                 aes(lon, lat),
                 binwidth = 1) +
      scale_fill_viridis_c(direction = -1) +
      coord_quickmap(expand = 0) +
      labs(
        title = paste("Random subsetting with latitudinal rarefaction of year", i_year),
        subtitle = paste("Nr of observations", nrow(resample_lat_year)),
        x = "Longitude",
        y = "Latitude"
      )
  )

}

```

```{r standard_resample_distribution_maps}

# read in random model sampling file
standard <-
  read_csv(paste(path_preprocessing,
                 "GLODAPv2.2020_preprocessed_model_runA_final.csv",
                 sep = ""))

# plot random sampling cmorized grids in each year
#i_year <- "2000"
years <- c("1982", "1990", "2000", "2010", "2019")
for (i_year in years) {
  # i_year <- years[1]
  
  standard_year <- standard %>%
    filter(year == i_year)
  
  print(
    map +
      geom_bin2d(data = standard_year,
                 aes(lon, lat),
                 binwidth = 1) +
      scale_fill_viridis_c(direction = -1) +
      coord_quickmap(expand = 0) +
      labs(
        title = paste("Standard subsetting"),
        subtitle = paste("Nr of observations", nrow(standard_year)),
        x = "Longitude",
        y = "Latitude"
      )
  )

}

```

# GLODAP observations vs cmorized subset comparison

```{r read_GLODAP_with_cmorized_subset_file}

GLODAP_cmorized <- read_csv(paste(path_preprocessing,
                                  "GLODAPv2.2020_preprocessed_model_runA_both.csv",
                                  sep = ""))

```

```{r GLODAP_vs_cmorized_subset_comparison}

# for loop across variables
obs_var <-
  c(
    #"tco2"
    # "talk",
    # "oxygen",
    # "aou",
    # "nitrate",
    # "phosphate",
    # "silicate",
    # "sal",
    # "temp",
    # "theta",
     "gamma"
  )

model_var <-
  c(
    #"dissic_model"
    # "talk_model",
    # "o2_model",
    # "aou_model",
    # "no3_model",
    # "po4_model",
    # "si_model",
    # "so_model",
    # "temp_model",
    # "thetao_model",
     "gamma_model"
  )

for (i in 1:11) {
  i <- 1
  
  # select correlated observation and model variable
  GLODAP_cmorized_var <- GLODAP_cmorized %>%
    select(year,
           month,
           lat,
           lon,
           depth,
           basin_AIP,
           !!sym(obs_var[i]),
           !!sym(model_var[i])) %>%
    drop_na() %>%
    mutate(
      season = case_when(
        month %in% c(3, 4, 5) ~ "Spring",
        month %in% c(6, 7, 8) ~ "Summer",
        month %in% c(9, 10, 11) ~ "Autumn",
        month %in% c(12, 1, 2) ~ "Winter"
      )
    )
  
  GLODAP_cmorized_var$season_f = factor(GLODAP_cmorized_var$season, levels=c("Spring", "Summer", "Autumn", "Winter"))
  
  # calculate equal axis limits and binwidth
  axis_lims <- GLODAP_cmorized_var %>%
    summarise(max_value = max(c(max(!!sym(
      obs_var[i]
    )),
    max(
      !!sym(model_var[i])
    ))),
    min_value = min(c(min(!!sym(
      obs_var[i]
    )),
    min(
      !!sym(model_var[i])
    ))))
  
  binwidth_value <- (axis_lims$max_value - axis_lims$min_value) / 40
  axis_lims <- c(axis_lims$min_value, axis_lims$max_value)
  
  # obs-model plot (season)
a <- ggplot(GLODAP_cmorized_var, aes(
      x = !!sym(obs_var[i]),
      y = !!sym(model_var[i])
    )) +
      geom_bin2d(binwidth = binwidth_value) +
      labs(subtitle = "Seasonal comparison",
           x = expression(Observed~DIC ~ (mu * mol ~ kg ^ "-1")),
           y = expression(Simulated~DIC ~ (mu * mol ~ kg ^ "-1"))) +
      scale_fill_viridis_c(trans = "log10") +
      geom_abline(slope = 1,
                  col = 'red') +
      coord_equal(xlim = axis_lims,
                  ylim = axis_lims) +
      facet_wrap( ~ season_f)
  
  # obs-model plot (year)
ggplot(GLODAP_cmorized_var, aes(
      x = !!sym(obs_var[i]),
      y = !!sym(model_var[i])
    )) +
      geom_bin2d(binwidth = binwidth_value) +
      labs(subtitle = "All years comparison",
           x = expression(Observed~DIC ~ (mu * mol ~ kg ^ "-1")),
           y = expression(Simulated~DIC ~ (mu * mol ~ kg ^ "-1"))) +
      scale_fill_viridis_c(trans = "log10") +
      geom_abline(slope = 1, col = 'red') +
      coord_equal(xlim = axis_lims,
                  ylim = axis_lims)
  
  # Calculate variable offset
  GLODAP_cmorized_var <- GLODAP_cmorized_var %>%
    mutate(offset = !!sym(model_var[i]) -!!sym(obs_var[i]))
  
  # Calculate annual mean offset
  GLODAP_cmorized_var_year <- GLODAP_cmorized_var %>%
    group_by(year) %>%
    summarise(offset = mean(offset)) %>%
    ungroup()
  
  # plot annual mean offset
  print(
    GLODAP_cmorized_var_year %>%
      ggplot(aes(year, offset)) +
      geom_point() +
      geom_line() +
      labs(title = "Annual mean offset",
           subtitle = paste(sym(model_var[i]), "-", sym(obs_var[i])))
  )
  
  # zonal mean section of the model-observations offset for each basin
  for (i_basin_AIP in unique(GLODAP_cmorized_var$basin_AIP)) {
    i_basin_AIP = "Atlantic"
    GLODAP_cmorized_var_zonal <- GLODAP_cmorized_var %>%
      filter(basin_AIP == i_basin_AIP) %>%
      group_by(lat, depth) %>%
      summarise(offset = mean(offset, na.rm=TRUE)) %>%
      ungroup()
    
    # plot zonal mean section (model - obs)
    lat_max <- params_global$lat_max
    lat_min <- params_global$lat_min
    limit <-
      quantile(abs(GLODAP_cmorized_var_zonal$offset), 0.98) * c(-1, 1)
    
    print(
      GLODAP_cmorized_var_zonal %>%
        ggplot(aes(lat, depth, z = offset)) +
        stat_summary_2d(binwidth = c(1,100)) +
        scale_fill_scico(
          palette = "vik",
          name = expression(gamma~offset),
          # name = expression(atop(DIC~offset,
          #                                 (mu * mol ~ kg ^ {-1}))),
          limit = limit,
          na.value = "red"
        ) +
        coord_cartesian(expand = 0,
                        xlim = c(lat_min, lat_max)) +
        scale_y_reverse() +
        labs(
          subtitle = paste("(d) Basin:",
                           i_basin_AIP),
          x = "Latitude",
          y = "Depth"
        )
    )
  }
}

```

