---
title: "Random subset variable A"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r define_paths}
path_GLODAP_preprocessing <-
  paste(path_root, "/observations/preprocessing/", sep = "")
path_cmorized <-
  "/nfs/kryo/work/loher/CESM_output/RECCAP2/cmorized_Dec2020/split_monthly/"
path_preprocessing  <-
  paste(path_root, "/model/preprocessing/", sep = "")
```

```{r load_libraries_specific, include=FALSE}
library(stars)
library(lubridate)
library(rqdatatable)
library(gsw)
library(marelac)
```


# Randomly subset model data

Here we randomly subset cmorized (1x1) model with variable forcing, according to the total number of GLODAP observations for the whole period from a previously cleaned file. The number for the annual subset remains the same for each year, which could be expressed by the total number of observations divided by number of years.

Besides, Model results are given in [mol m^-3^], whereas GLODAP data are in [µmol kg^-1^]. This refers to the variables:

- DIC  
- ALK  
- O2  
- NO3  
- PO4  
- SiO3  
- AOU (calculated)

For comparison, model results were converted from [mol m^-3^] to [µmol kg^-1^]

# Spatial distribution

```{r random_resample_distribution_maps}

# read in random model sampling file
resample_lat <-
  read_csv(paste(path_preprocessing,
                 "GLODAPv2.2020_preprocessed_model_runA_random_subset_lat.csv",
                 sep = ""))

# plot random sampling cmorized grids in each year
#i_year <- "2000"
years <- c("1982", "1990", "2000", "2010", "2019")
for (i_year in years) {
  # i_year <- years[1]
  
  resample_lat_year <- resample_lat %>%
    filter(year == i_year)
  
  print(
    map +
      geom_bin2d(data = resample_lat_year,
                 aes(lon, lat),
                 binwidth = 1) +
      scale_fill_viridis_c(direction = -1) +
      coord_quickmap(expand = 0) +
      labs(
        title = paste("Random subsetting with latitudinal rarefaction of year", i_year),
        subtitle = paste("Nr of observations", nrow(resample_lat_year)),
        x = "Longitude",
        y = "Latitude"
      )
  )

}

```

```{r standard_resample_distribution_maps}

# read in random model sampling file
standard <-
  read_csv(paste(path_preprocessing,
                 "GLODAPv2.2020_preprocessed_model_runA_final.csv",
                 sep = ""))

# plot random sampling cmorized grids in each year
#i_year <- "2000"
years <- c("1982", "1990", "2000", "2010", "2019")
for (i_year in years) {
  # i_year <- years[1]
  
  standard_year <- standard %>%
    filter(year == i_year)
  
  print(
    map +
      geom_bin2d(data = standard_year,
                 aes(lon, lat),
                 binwidth = 1) +
      scale_fill_viridis_c(direction = -1) +
      coord_quickmap(expand = 0) +
      labs(
        title = paste("Standard subsetting"),
        subtitle = paste("Nr of observations", nrow(standard_year)),
        x = "Longitude",
        y = "Latitude"
      )
  )

}

```

